Summary of the things that I learned

### 2024-05-01 report/chapters/algebraic-structures.tex
Start working on Hyland's Taylor fibration

>>>>

\subsection{The initial presheaf}
\begin{example}
  Let $ T $ be an algebraic theory. We can construct a $ T $-presheaf $ P $, with $ P_n = \emptyset $. Then $ \bullet : P_m \times T_n^m \to P_m $ is trivial, and the presheaf laws hold trivially.
\end{example}

\begin{lemma}
  This is indeed the initial presheaf.
\end{lemma}
\begin{proof}
  Let $ Q $ be a $ T $-presheaf. For all $ n $, since $ P_n $ is empty, there is only one possible function $ P_n \to Q_n $. These functions trivially satisfy the presheaf morphism laws, so they constitute the unique presheaf morphism $ P \to Q $.
\end{proof}

### 2024-05-01 report/chapters/category-theory.tex
Start working on Hyland's Taylor fibration

Now, there is also another (classically equivalent) definition of the category of retracts:
\begin{definition}
  Let $ Y: C \to P C $ be the Yoneda embedding of $ C $ into its category of presheaves. We define the category $ \hat \R $ as the full subcategory of $ P C $ consisting of objects $ F : P C $ such that there exists an object $ c: C $ and a retraction of $ Y(c) $ onto $ F $. Note that the existence of $ c $ and the retraction is \textit{mere existence} (see the chapter on Univalent Foundations) because we need this to be a subcategory of $ P C $.
\end{definition}

\begin{lemma}
  $ \hat \R $ is the Rezk completion of $ \R $.
\end{lemma}
\begin{proof}
  First of all, note that $ \hat \R $ is univalent, since it is a presheaf category (see Theorem 4.5 in \cite{univalent-categories}). So we must show that we have a weak equivalence from $ \R $ to $ \hat \R $.

  Secondly, by Corollary \ref{cor:karoubi-presheaf} we have an adjoint equivalence $ {\op{\iota_C}}_*: P \R \xrightarrow \sim P C $ and by Lemma \ref{lem:Yoneda-restriction-commutes}, the following diagram 2-commutes:
  \begin{center}
    \begin{tikzcd}
      C \arrow[r, "\iota_C"] \arrow[d, "Y"] & \R \arrow[d, "Y"]\\
      P C & P \R \arrow[l, "\sim"', "{\op{\iota_C}}_*"]
    \end{tikzcd}
  \end{center}

  By Lemma \ref{lem:karoubi-is-retract}, every object of $ \R $ is a retract of some object in $ c $. Therefore, every object in the image of $ \R $ inside $ P C $ is a retract of some object in the image of $ C $, so we can turn $ Y \bullet {\op{\iota_C}}_* : \R \to P C $ into a fully faithful embedding into $ \hat \R $.

  Now, to show that it is essentially surjective, take an object $ F : \hat \R $. That is, take a retract \begin{tikzcd} F \arrow[r, shift left, "s"] & Y(c) \arrow[l, shift left, "r"] \end{tikzcd}. Note that $ r \cdot s $ is an idempotent on $ Y(c) $. Since $ Y $ is a fully faithful embedding, it gives an equivalence on the hom-sets, so we have an element $ e : C(c, c) $ that maps to $ r \cdot s : P C (Y(c), Y(c)) $. We will show that $ (c, e) $ is our preimage in $ \R $ of $ F $.

@@

\begin{remark}
  As shown above, $ \R $ and $ \hat \R $ have a very strong relation. In classical mathematics, they are even equivalent. However, since $ \R $ is not univalent, we do not have a full equivalence in univalent foundations and our choice of definition will have consequences. One one hand, $ \hat \R $ is univalent. On the other hand, the objects of $ \R $ are uniquely associated with one idempotent morphism in $ C $.

  In this thesis, I will choose to interpret the `category of retracts' to mean $ \R $. This is because Dana Scott and Paul Taylor perform most of their constructions explicitly using the (idempotent) morphisms of $ C $, which is closest to working in $ \R $.

  It is however very well possible that most of the proofs can be made to work in $ \hat \R $ as well. For future research, it would be interesting to see what the differences are between working in $ \R $ and $ \hat \R $ for these proofs.
\end{remark}

>>>>

Now, there is also another (classically equivalent) definition of the category of retracts:
\begin{definition}\label{def:karoubi'}
  Let $ Y: C \to P C $ be the Yoneda embedding of $ C $ into its category of presheaves. We define the category $ \hat C $ as the full subcategory of $ P C $ consisting of objects $ F : P C $ such that there exists an object $ c: C $ and a retraction of $ Y(c) $ onto $ F $. Note that the existence of $ c $ and the retraction is \textit{mere existence} (see the chapter on Univalent Foundations) because we need this to be a subcategory of $ P C $.
\end{definition}

\begin{lemma}\label{lem:retracts-rezk}
  $ \hat C $ is the Rezk completion of $ \overline C $.
\end{lemma}
\begin{proof}
  First of all, note that $ \hat C $ is univalent, since it is a presheaf category (see Theorem 4.5 in \cite{univalent-categories}). So we must show that we have a weak equivalence from $ \overline C $ to $ \hat C $.

  Secondly, by Corollary \ref{cor:karoubi-presheaf} we have an adjoint equivalence $ {\op{\iota_C}}_*: P \overline C \xrightarrow \sim P C $ and by Lemma \ref{lem:Yoneda-restriction-commutes}, the following diagram 2-commutes:
  \begin{center}
    \begin{tikzcd}
      C \arrow[r, "\iota_C"] \arrow[d, "Y"] & \overline C \arrow[d, "Y"]\\
      P C & P \overline C \arrow[l, "\sim"', "{\op{\iota_C}}_*"]
    \end{tikzcd}
  \end{center}

  By Lemma \ref{lem:karoubi-is-retract}, every object of $ \overline C $ is a retract of some object in $ c $. Therefore, every object in the image of $ \overline C $ inside $ P C $ is a retract of some object in the image of $ C $, so we can turn $ Y \bullet {\op{\iota_C}}_* : \overline C \to P C $ into a fully faithful embedding into $ \hat C $.

  Now, to show that it is essentially surjective, take an object $ F : \hat C $. That is, take a retract \begin{tikzcd} F \arrow[r, shift left, "s"] & Y(c) \arrow[l, shift left, "r"] \end{tikzcd}. Note that $ r \cdot s $ is an idempotent on $ Y(c) $. Since $ Y $ is a fully faithful embedding, it gives an equivalence on the hom-sets, so we have an element $ e : C(c, c) $ that maps to $ r \cdot s : P C (Y(c), Y(c)) $. We will show that $ (c, e) $ is our preimage in $ \overline C $ of $ F $.

@@

\begin{remark}
  As shown above, $ \overline C $ and $ \hat C $ have a very strong relation. In classical mathematics, they are even equivalent. However, since $ \overline C $ is not univalent, we do not have a full equivalence in univalent foundations and our choice of definition will have consequences. One one hand, $ \hat C $ is univalent. On the other hand, every object of $ \overline C $ is uniquely associated with one idempotent morphism in $ C $.

  In this thesis, I will choose to interpret the `category of retracts' to mean $ \overline C $. This is because Dana Scott and Paul Taylor perform most of their constructions explicitly using the (idempotent) morphisms of $ C $, which is closest to working in $ \overline C $.

  It is however very well possible that most of the proofs can be made to work in $ \hat C $ as well. For future research, it would be interesting to see what the differences are between working in $ \overline C $ and $ \hat C $ for these proofs.
\end{remark}

### 2024-05-01 report/chapters/paper.tex
Start working on Hyland's Taylor fibration

\begin{lemma}
  We have a natural isomorphism between the two functors from $ \R $ to $ P \mathbf L $ in the diagram above.
\end{lemma}
\begin{proof}
  Note:
  \[ (Y \bullet j_*)(A)(n) = (j \bullet (Y(A)))(n) = Y(A)(j(n)) = \R(U^n, A) \]
  \[ (\iota \bullet \sim)(A)(n) = \{ a : L_n \mid A a = a \} \]

  By Scott's representation theorem, we have $ \mathbb R(U^n, U) \cong L_n, $ via
  \[ f \mapsto \iota_{0, n}(f) ((((c, x_1), x_2), \dots), x_n). \]
  Restricting this equivalence to the morphisms to $ A $ yields
  \[ \R(U^n, A) \cong \{ a : L_n \mid A a = a \}. \]
  \TODO: Check coherence.
\end{proof}

@@

Recall that Paul Taylor constructed a class of display maps for $ \R $.

>>>>

\begin{lemma}
  The square in the diagram above 2-commutes.
\end{lemma}
\begin{proof}
  Note:
  \[ (Y \bullet j_*)(A)(n) = (j \bullet (Y(A)))(n) = Y(A)(j(n)) = \R(U^n, A) \]
  \[ (\iota \bullet \sim)(A)(n) = \{ a : L_n \mid A a = a \} \]

  By Scott's representation theorem, we have $ \mathbb R(U^n, U) \cong L_n, $ via
  \[ f \mapsto \iota_{0, n}(f) ((((c, x_1), x_2), \dots), x_n). \]
  Restricting this equivalence to the morphisms to $ A $ yields
  \[ \R(U^n, A) \cong \{ a : L_n \mid A a = a \}. \]
  Of course, this is natural in $ A $, since for $ g: \R(A, B) $, postcomposing elements of $ \R(U^n, A) $ by $ g $ is equivalent to applying $ g $ to the elements of $ \{ a : L_n \mid A a = a \} $.

  \TODO: Check naturality in $ n $.
\end{proof}

@@

Recall that Paul Taylor constructed a category of `dependent retracts' $ \R^A $, and then defined a class of display maps for $ \R $ in such a way that it gives a weak equivalence $ \sum_A : \R^A \xrightarrow{\sim} (\R \downarrow_D A) $. I.e., a map $ f: \R(X, A) $ is defined to be a display map if it is isomorphic to something in the image of the functor $ \sum_A $. We saw that these display maps can also be characterized as the retracts of $ p_1: A \times U \to A $ in $ (\R \downarrow A) $.

Now, noticing that we can embed $ \R $ as a full subcategory in $ \Pshf L $, Hyland attempts to do a similar thing for $ \Pshf L $. Noting that the universal object $ U : \R $ is embedded as the theory presheaf $ L : \Pshf L $, he defines a full subcategory $ \R(A) \subseteq (\Pshf L \downarrow A) $, consisting precisely of the retracts of $ p_1: A \times L \to A $ in $ (\Pshf L \downarrow A) $.

\begin{remark}
  Again, being a retract of $ p_1 $ is not a mere proposition. For example, consider $ \id L : D(L, L) $:
  \begin{center}
    \begin{tikzcd}
      L \arrow[rr, shift left, "s"] \arrow[rd, "\id L"'] & & L \times L \arrow[ll, shift left, "r"] \arrow[ld, "p_1"]\\
      & L
    \end{tikzcd}
  \end{center}
  There are multiple possible retraction-section pairs $ (r, s) $ that make the diagram commute. For example, we can take $ s $ to be the diagonal embedding $ \Delta : L \to L \times L $ (sending $ l $ to $ (l, l) $). Then we can take $ r = p_1 $ or $ r = p_2 $. We can also take $ r = \pi_1 $ and $ s = \langle \id L, g \rangle $ the product morphism of $ \id L $ and any (possibly constant) endomorphism $ g: L \to L $.

  Therefore, we need to take the propositional truncation of the existence of $ r $ and $ s $.
\end{remark}

\begin{remark}
  Taylor would write $ (\Pshf L \downarrow_D A) $ instead of $ \R(A) $, with $ D(X, A) $ the retracts $ f: X \to A $ of $ p_1 : L \times A \to A $. However, note that $ D $ is not a class of display maps. For example, for $ D $ to be a class of display maps, all terminal projections $ ! : X \to T $ need to be in $ D $. Since for any category $ C $ with terminal object $ T $, we have an equivalence $ (C \downarrow T) \cong C $, having all terminal projections in $ D $ is equivalent to every object in $ \Pshf L $ being a retract of $ L $.

  However, if we take $ X $ to be the empty (initial) presheaf, we see that there are no morphisms $ L \to X $, because the $ L_n $ are nonempty and the $ X_n $ are empty.

  There also is a more category-theoretical argument: Suppose that every object of $ \Pshf L $ is a retract of $ L $. Via the equivalence $ \Pshf L \cong L_1 $, we see that every object of $ P L_1 $ is a retract of $ Y(\star) $. However then the embedding of $ \hat \R $ (Definition \ref{def:karoubi'}) into $ P L_1 $ is an adjoint equivalence. Since $ Y_{\R} \bullet {\op{\iota_{L_1}}}_* : \R \to \hat \R $ and $ {\op{\iota_{L_1}}}_* : P \R \to P L_1 $ are weak equivalences, (Lemma \ref{lem:retracts-rezk}, Corollary \ref{cor:karoubi-presheaf}), $ Y_\R : \R \to P \R $ is a weak equivalence as well, but the Yoneda embedding is never essentially surjective. For example, the constant empty presheaf $ E: c \mapsto \emptyset $ is not representable, since for all $ c : C $, $ \id c = Y(c)(c) $, so $ Y(c) $ is not isomorphic to $ E $.
\end{remark}

### 2024-05-01 transcript/summary.tex
Start working on Hyland's Taylor fibration

  \section{Cartesian multicategory}
  Yet another way to think of a cartesian operad is as a one-object cartesian multicategory. A \textbf{cartesian multicategory} is a multicategory with an $ S_n $-action on the hom-sets (in other words: we can permute the arguments of the morphsims) and duplication/diagonal/contraction operations
  \[ \mathrm{Hom}(c_1, \dots, c_k, c_k, \dots, c_n; c) \to \mathrm{Hom}(c_1, \dots, c_k, \dots, c_n; c) \]
  and deletion/projection/weakening operations
  \[ \mathrm{Hom}(c_1, \dots, c_k, \dots, c_n; c) \to \mathrm{Hom}(c_1, \dots, c_{k-1}, c_{k+1}, \dots, c_n; c) \]

>>>>

  \section{Cartesian multicategory}
  Yet another way to think of a cartesian operad is as a one-object cartesian multicategory. A \textbf{cartesian multicategory} is a multicategory with an $ S_n $-action on the hom-sets (in other words: we can permute the arguments of the morphisms) and duplication/diagonal/contraction operations
  \[ \mathrm{Hom}(c_1, \dots, c_k, c_k, \dots, c_n; c) \to \mathrm{Hom}(c_1, \dots, c_k, \dots, c_n; c) \]
  and deletion/projection/weakening operations
  \[ \mathrm{Hom}(c_1, \dots, c_k, \dots, c_n; c) \to \mathrm{Hom}(c_1, \dots, c_{k-1}, c_{k+1}, \dots, c_n; c) \]
